
\newgeometry{total={210mm,297mm},left=20mm,right=20mm,bindingoffset=5mm, top=25mm,bottom=25mm} 
\begin{partwithabstract}{Results \& Experiments}
    \par{
    The objective of this modelling procedure is to procedure a model that can estimate segmentation masks for the 5 lumbar vertebrae from \acrfull{ct} and \acrfull{mri} scans of human patients, 
    based on point annotation of these 6 classes ($L1$ to $L5$ and 1 background class).
}
\par{
    First, as described in chapter \ref{sec:reference_model}, a reference is calculated with which to compare the performance of the weakly supervised models.
    The reference models have the same architecture and are trained with full annotation on the same dataset as the weakly supervised models. 
}
\par{
    The first step in the modelling procedure based on point annotation masks is the construction of \textit{single-dimension} models.
    Single-dimension indicates that these models investigate the performance of a model that outputs 2D segmentation masks based on 2D input images.
    By slicing the scan volumes along one of three central dimensional axes, a stack of two-dimensional images is created together with the point annotation maps extracted from the segmentation masks of these images.
    Chapter \ref{sec:singleDimension} presents the results of several experiments. 
    Different model hyperparameters are tested.
    Models with different loss functions are trained on point annotation maps where the number of annotation points extracted from the full masks is varied.
}
\par{
    the second step in the modelling procedure is the combination of the single-dimension models to estimate pseudo-masks.
    These pseudo masks can finally be used to train a model as pseudo full annotation masks.
    Chapter \ref{sec:combination} first discusses the technique used to combine single-dimension models and then concludes with the results obtained by training a final model on the pseudo masks resulting from the combination.
}
\end{partwithabstract}
\restoregeometry



\input{/home/thesis/3_Results/result_reference}

\input{/home/thesis/3_Results/result_singleDimension}

\input{/home/thesis/3_Results/result_Combination}