\chapter*{Foreword}
\addcontentsline{toc}{chapter}{Foreword}

\par{
Before you lies my master thesis, "Spine vertebrae segmentation in 3D CT scan images".
This document presents my investigation of the weakly supervised segmentation problem applied on medical scans of the human lumbar spine.
I hope to convince the reader of the usefulness of the two loss function components I devised for this problem and the algorithm I developed to combine three segmentation models of two-dimensional slices of the volumes to produce an improved segmentation result.
}
\par{
I have developed this research question with my supervisors, Dr J. Roels and Dr B. Vankeirsbilck, who encouraged me to orient my thesis towards weakly supervised problems.
}
\par{
I want to thank my supervisors for their guidance and support during researching this topic and developing the procedure described in this document.
All the more, given the extraordinary conditions we have all experienced in the past two years.
}
\par{
I also want to thank my colleagues and coordinators at Verhaert for their moral support and understanding during the last two years.
The combination of a job and academic studies is a demanding undertaking.
I was lucky to experience the support and understanding I did.
}
\par{
My parents deserve a particular note of thanks for their support and kind words. These have, as always, served me well.
}\newline
\vspace{1cm}
I hope you enjoy the reading,\newline
Jan Alexander\newline
\vspace{1cm}
Gent, August 2021
\clearpage\null\newpage